---
title: "Day 1 - Session 1"
author: "Melinda Higgins"
date: "7/28/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abalone Dataset

For this exercise we will be working with the `abalone` dataset from the University of California - Irvine Machine Learning Repository. This dataset can be downloaded from [https://archive.ics.uci.edu/ml/datasets/abalone](https://archive.ics.uci.edu/ml/datasets/abalone).

The dataset has 9 measurements on 4177 abalones of different ages and sexes. The data comes from this study: Warwick J. Nash, Tracy L. Sellers, Simon R. Talbot, Andrew J. Cawthorn and Wes B. Ford (1994) "The Population Biology of Abalone (_Haliotis_ species) in Tasmania. I. Blacklip Abalone (_H. rubra_) from the North Coast and Islands of Bass Strait", Sea Fisheries Division, Technical Report No. 48 (ISSN 1034-3288).

The purpose of the study was to predicting the age of abalone from physical measurements. The age of abalone is determined by cutting the shell through the cone, staining it, and counting the number of rings through a microscope -- a boring and time-consuming task. This process also kills the abalones! Other measurements, which are easier to obtain, can be used to predict the age. Further information, such as weather patterns and location (hence food availability) may be required to solve the problem.

## Load the abalone dataset

For this R project, the abalone data have already been downloaded as a CSV file. We will load this data file from the R project directory into "memory" for use in this Rmarkdown file with our R analysis codes.

```{r}
# load the abalone dataset
library(readr)
abalone <- readr::read_csv("abalone.csv")
```

## Display the top 6 rows of the dataset

Use the base R function `head()` to show the top 6 rows of the dataset. To learn more about the `head()` function, type `head` in the help window or in the console type `help(head)` or `?head` to learn more.

```{r}
head(abalone)
```

## View a tibble of the dataset

```{r}
library(dplyr)
tibble(abalone)
```

## Get a glimpse of the dataset

```{r}
glimpse(abalone)
```

## Arrange (sort) data by height, view top 6 rows

```{r}
abalone %>%
  arrange(height) %>%
  head()
```

## Arrange (sort) data by rings in descending order, view top 10 rows

```{r}
abalone %>%
  arrange(desc(rings)) %>%
  head(10)
```

## Arrange by rings, view bottom 5 rows

```{r}
abalone %>%
  arrange(rings) %>%
  tail(5)
```

